Reduced plasticity and opportunity for selection in terrestrial ectotherm populations under climate change

### 2023-01-18 ms.rmd
Noted in stats methods

We first tested the degree to which acute and acclimation $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ effects varied by habitat type (i.e., terrestrial, freshwater, and marine ecosystems). Models included an interaction between effect type (i.e., acute or acclimation) and habitat. Reduced mean $lnRR_{Q_{10}}_{acclimation}$ relative to $lnRR_{Q_{10}}_{acute}$ indicates that acclimation to thermal environments results in partial compensation of physiological rates (i.e., phenotypic plasticity), whereas no differences between $lnRR_{Q_{10}}_{acute}$ and $lnRR_{Q_{10}}_{acclimation}$ suggests organisms are not capable of physiological plasticity [@Havird2020]. In contrast, a difference in $lnCVR_{Q_{10}}_{acclimation}$ relative to $lnCVR_{Q_{10}}_{acute}$ would imply that changes in between individual variation in physiological rates across 10$^{\circ}$C differ depending on whether acute or acclimation responses are measured. If the interaction between effect type and habitat was not supported, then we fit a model that only contained additive effects of effect type and habitat. Following on from these models, we subset each habitat type and explored how mean $lnRR_{Q_{10}}$ changed across traits. Within each habitat (marine, freshwater, and terrestrial) we fit a series of models that included an interaction between effect type (acute / acclimation) and trait category (as defined above). Variance in effects within trait categories appeared to vary depending on the trait type in question. Comparison of a model with and without heteroscedastic residual variance favored a model with heteroscedastic residual variance across trait categories ($\Delta AIC_{c}$; marine = 58, freshwater = 120, and terrestrial = 25)

@@

    # Filter out antioxidant and NAs. For terrestrial, need to filter out endurance and max_MR because only 3 effect sizes
  lnRRQ10_data_m <- lnRRQ10_data_m %>% filter(!trait_category =="antiox") %>% filter(!is.na(c_accl_time))
  lnRRQ10_data_f <- lnRRQ10_data_f %>% filter(!is.na(c_accl_time))
  lnRRQ10_data_t <- lnRRQ10_data_t %>% filter(!is.na(c_accl_time)) %>% filter(!trait_category %in% c("endurance", "max_MR"))

@@

  # Bayesian model
   model1.3_t_bayes <- brms::brm(effect_size | se(sqrt(V)) ~  c_accl_time + type + trait_category + type:trait_category + (~1 | record_num) + (1 | gr(species_full, cov = A_t)) + (1 | spp) + (1 | trait) + (1 + trait_category | gr(obs, cor = FALSE)), data2 = list(A_t=A_t), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 4, thin = 5, iter = 5000, warmup = 1000, cores = 4, data = lnRRQ10_data_t)

>>>>

We first tested the degree to which acute and acclimation $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ effects varied by habitat type (i.e., terrestrial, freshwater, and marine ecosystems). Models included an interaction between effect type (i.e., acute or acclimation) and habitat. Reduced mean $lnRR_{Q_{10}}_{acclimation}$ relative to $lnRR_{Q_{10}}_{acute}$ indicates that acclimation to thermal environments results in partial compensation of physiological rates (i.e., phenotypic plasticity), whereas no differences between $lnRR_{Q_{10}}_{acute}$ and $lnRR_{Q_{10}}_{acclimation}$ suggests organisms are not capable of physiological plasticity [@Havird2020]. In contrast, a difference in $lnCVR_{Q_{10}}_{acclimation}$ relative to $lnCVR_{Q_{10}}_{acute}$ would imply that changes in between individual variation in physiological rates across 10$^{\circ}$C differ depending on whether acute or acclimation responses are measured. If the interaction between effect type and habitat was not supported, then we fit a model that only contained additive effects of effect type and habitat. Following on from these models, we subset each habitat type and explored how mean $lnRR_{Q_{10}}$ changed across traits. Within each habitat (marine, freshwater, and terrestrial) we fit a series of models that included an interaction between effect type (acute / acclimation) and trait category (as defined above). Variance in effects within trait categories appeared to vary depending on the trait type in question. Comparison of a model with and without heteroscedastic residual variance favored a model with heteroscedastic residual variance across trait categories ($\Delta AIC_{c}$; marine = 58, freshwater = 120, and terrestrial = 12). To ensure models converged we limited to trait categories for each habitat with six or more effect sizes.

@@

    # Filter out antioxidant and NAs. For terrestrial, need to filter out endurance and max_MR because only 3 effect sizes
  lnRRQ10_data_m <- lnRRQ10_data_m %>% filter(!trait_category =="antiox") %>% filter(!is.na(c_accl_time))
  lnRRQ10_data_f <- lnRRQ10_data_f %>% filter(!is.na(c_accl_time))
  lnRRQ10_data_t <- lnRRQ10_data_t %>% filter(!is.na(c_accl_time)) %>% filter(!trait_category %in% c("endurance", "max_MR", "other"))

@@

  # Bayesian model
   model1.3_t_bayes <- brms::brm(effect_size | se(sqrt(V)) ~  c_accl_time + type + trait_category + type:trait_category + (~1 | record_num) + (1 | gr(species_full, cov = A_t)) + (1 | spp) + (1 | trait) + (1 + trait_category | gr(obs, cor = FALSE)), data2 = list(A_t=A_t), family = gaussian(), prior = c(prior(normal(0, 10), "Intercept"), prior(student_t(3, 0, 10), "sd")), control = list(adapt_delta = 0.98, max_treedepth = 12), chains = 8, thin = 5, iter = 3000, warmup = 1000, cores = 8, data = lnRRQ10_data_t)
